---
title: "Primers-ggplot2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Primers ggplot2

```{r}
library(tidyverse)
```


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut), width = 1)
```

```{r}
ggplot(data=mpg)+ geom_bar(aes(x=fct_infreq(class), fill=class, alpha=0.5))
```


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity))
```


```{r}
#para ver en forma de histograma
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

```{r}
#stacked bars
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "stack")
```

```{r}
#barras apiladas
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```



```{r}
geom_bar(mapping = aes(x = cut, fill = clarity), position = "identity")
```

## Ejercicio 6 (recrear el grafico)

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = clarity), position = "fill", width=1)
```


## Ejercicio 7 (recrear el grafico)

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = cut), position = "dodge")
```

# facet_grid()

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color)) +
  facet_grid(clarity ~ cut)
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color)) +
  facet_grid(clarity ~ .)
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color)) +
  facet_grid(.~cut)
```

# facet_wrap()


```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = cut))+
  facet_wrap( ~ cut)

```

## scales

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = color, fill = cut)) +
  facet_wrap( ~ cut, scales = "free_y")
```


In this tutorial, you learned how to make bar charts; but much of what you learned applies to other types of charts as well. Here’s what you should know:

1. Bar charts are the basis for histograms, which means that you can interpret histograms in a similar way.
2. Bars are not the only geom in ggplot2 that use the fill aesthetic. You can use both fill and color aesthetics with any geom that has an “interior” region.
3. You can use the same position adjustments with any ggplot2 geom: "identity", "stack", "dodge", "fill", "nudge", and "jitter" (we’ll learn about "nudge" and "jitter" later). Each geom comes with its own sensible default.
4. You can facet any ggplot2 plot by adding facet_grid() or facet_wrap() to the plot call.

Bar charts are an excellent way to display the distribution of a categorical variable. In the next tutorial, we’ll meet a set of geoms that display the distribution of a continuous variable. 




# Histogramas

```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = carat))
```

```{r}
ggplot(data=diamonds) + geom_histogram(aes(x=carat), binwidth=0.5)
```

"Good job! The most common diamond size is about 0.5 carats. Larger sizes become progressively less frequent as carat size increases. This accords with general knowledge about diamonds, so you may be prompted to stop exploring the distribution of carat size. But should you?"

#Vemos nuevos patrones al cambiar de bindwidth
```{r}
ggplot(data=diamonds) + geom_histogram(aes(x=carat), binwidth=0.5)

```

"Good job! The new binwidth reveals a new phenomena: carat sizes like 0.5, 0.75, 1, 1.5, and 2 are much more common than carat sizes that do not fall near a common fraction. Why might this be?"

```{r}
ggplot(data=diamonds) + geom_histogram(aes(x=carat), binwidth=0.01)

```

"Good job! The new binwidth reveals another phenomena: each peak is very right skewed. In other words, diamonds that are 1.01 carats are much more common than diamonds that are .99 carats. Why would that be?"



## Recrear el siguiente grafico

```{r}
```


```{r}
ggplot(data=diamonds) + geom_histogram(aes(x=price, fill=cut, binwidth=1000, boundary=0))
```


But this display has some shortcomings:

* it is difficult to see the “shapes” of the individual distributions
* it is difficult to compare the individual distributions, because they do not share a common baseline value for y

A solution

We can improve the plot by using a different geom to display the distributions of price values. ggplot2 includes three geoms that display the same information as a histogram, but in different ways:

    geom_freqpoly()
    geom_density()
    geom_dotplot()

# frequency polygons

```{r}
ggplot(data = diamonds) +
  geom_freqpoly(mapping = aes(x = carat), binwidth = 0.25, boundary = 0)
```

```{r}

ggplot(data = diamonds) +
  geom_freqpoly(mapping = aes(x = carat, color=cut), binwidth = 0.25, boundary = 0)
```

geom_density()

Our frequency polygon highlights a second shortcoming with our graph: it is difficult to compare the shapes of the distributions because some sub-groups contain more diamonds than others. This compresses smaller subgroups into the bottom of the graph.


Density curves

geom_density() plots a kernel density estimate (i.e. a density curve) for each distribution. This is a smoothed representation of the data, analogous to a smoothed histogram.

Density curves do not plot count on the y
axis but density, which is analagous to the count divided by the total number of observations. Densities makes it easy to compare the distributions of sub-groups. When you plot multiple sub-groups, each density curve will contain the same area under its curve.

```{r}
ggplot(data = diamonds) +
  geom_density(mapping = aes(x = price, color = cut))
```


Density parameters

Density plots do not take bin, binwidth, and boundary parameters. Instead they recognize kernel and smoothing parameters that are used in the density fitting algorithm, which is fairly sophisticated.

In practice, you can obtain useful results quickly with the default parameters of geom_density(). If you’d like to learn more about density estimates and their tuning parameters, begin with the help page at ?geom_density().
geom_dotplot()

ggplot2 provides a final geom for displaying one dimensional distributions: geom_dotplot(). geom_dotplot() represents each observation with a dot and then stacks dots within bins to create the semblance of a histogram.

Dotplots can provide an intuitive display of the data, but they have several shortcomings. Dotplots are not ideal for large data sets like diamonds, and provide meaningless y
axis labels. I also find that the tuning parameters of geom_dotplot() make dotplots too slow to work with for EDA.

```{r}
ggplot(data = mpg) +
  geom_dotplot(mapping = aes(x = displ), dotsize = 0.5, stackdir = "up", stackratio = 1.1)

```

Exercise 9 - Facets

Instead of changing geoms, you can make the sub-groups in our original plot easier to compare by facetting the data. Extend the code below to facet by cut.

```{r}
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price, fill = cut), binwidth = 1000, boundary = 0)+
  facet_wrap(~cut)

```

Recap

In this tutorial, you learned how to visualize distributions with histograms, frequency polygons, and densities. But what should you look for in these visualizations?

* Look for places with lots of data. Tall bars reveal the most common values in your data; you can expect these values to be the “typical values” for your variable.

* Look for places with little data. Short bars reveal uncommon values. These values appear rarely and you might be able to figure out why.

* Look for outliers. Bars that appear away from the bulk of the data are outliers, special cases that may reveal unexpected insights.

* Sometimes outliers cannot be seen in a plot, but can be inferred from the range of the x axis. For example, many of the plots in this tutorial seemed to extend well past the end of the data. Why? Because the range was stretched to include outliers. When your data set is large, like diamonds, the bar that describes an outlier may be invisible (i.e. less than one pixel high).

* Look for clusters.

* Look for shape. The shape of a histogram can often indicate whether or not a variable behaves according to a known probability distribution.

The most important thing to remember about histograms, frequency polygons, and dotplots is to explore different binwidths. The binwidth of a histogram determines what information the histogram displays. You cannot predict ahead of time which binwidth will reveal unexpected information.



# Boxplots and counts

Boxplots display the relationship between a continuous variable and a categorical variable. Count plots display the relationship between two categorical variables. In this tutorial, you will learn how to use both. You will learn how to:

    Make and interpret boxplots
    Rotate boxplots by flipping the coordinate system of your plot
    Use violin plots and dotplots, two geoms that are similar to boxplots
    Make and interpret count plots

The tutorial is adapted from R for Data Science by Hadley Wickham and Garrett Grolemund, published by O’Reilly Media, Inc., 2016, ISBN: 9781491910399. You can purchase the book at shop.oreilly.com.

The tutorial uses the ggplot2 and dplyr packages, which have been pre-loaded for your convenience.


```{r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy))

```


Categorical and continuous

geom_boxplot() expects the y
axis to be continuous, but accepts categorical variables on the x axis. For example, here class is categorical. geom_boxplot() will automatically plot a separate boxplot for each value of x. This makes it easy to compare the distributions of points with different values of x.


## Recrear el grafico
```{r}
ggplot(data=diamonds)+ geom_boxplot(aes(x=cut, y=price))

```

You can change how outliers look in your boxplot with the parameters outlier.color, outlier.fill, outlier.shape, outlier.size, outlier.stroke, and outlier.alpha (outlier.shape takes a number from 1 to 25).

Unfortunately, you can’t tell geom_boxplot() to ignore outliers completely, but you can make outliers disappear by setting outlier.alpha = 0. Try it in the plot below.

```{r}
ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = cut, y = price), outlier.shape  = 24, 
               outlier.fill = "white", outlier.stroke = 0.25)
```


### ocultar outliers

```{r}
ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = cut, y = price), outlier.alpha = 0)
```

Aesthetics

Boxplots recognize the following aesthetics: alpha, color, fill, group, linetype, shape, size, and weight.

Of these group can be the most useful. Consider the plot below. It uses a continuous variable on the x
axis. As a result, geom_boxplot() is not sure how to split the data into categories: it lumps all of the data into a single boxplot. The result reveals little about the relationship between carat and price.



How to “cut” a continuous variable

ggplot2 provides three helper functions that you can use to split a continuous variable into categories. Each takes a continuous vector and returns a categorical vector that assigns each value to a group. For example, cut_interval() bins a vector into n equal length bins.

continuous_vector <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
continuous_vector

##  [1]  1  2  3  4  5  6  7  8  9 10

cut_interval(continuous_vector, n = 3)

##  [1] [1,4]  [1,4]  [1,4]  [1,4]  (4,7]  (4,7]  (4,7]  (7,10] (7,10] (7,10]
## Levels: [1,4] (4,7] (7,10]

The cut functions

The three cut functions are

    cut_interval() which makes n groups with equal range
    cut_number() which makes n groups with (approximately) equal numbers of observations
    cut_width() which makes groups with width width

Use one of three functions below to bin continuous_vector into groups of width = 2.


Exercise 4 - Apply a cut function

When you set the group
aesthetic of a boxplot, geom_boxplot() will draw a separate boxplot for each collection of observations that have the same value of whichever vector you map to group

.

This means we can split our carat plot by mapping group
to the output of a cut function, as in the code below. Study the code, then modify it to create a separate boxplot for each 0.25 wide interval of carat.


```{r}
ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = carat, y = price, group = cut_interval(carat, n = 2)))
```


```{r}
ggplot(data = diamonds) +
  geom_boxplot(mapping = aes(x = carat, y = price, group = cut_width(carat, width = 0.25)))
```



## coord_flip()
```{r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy))+
  coord_flip()
```

Similar Geoms
geom_dotplot()

Boxplots provide a quick way to represent a distribution, but they leave behind a lot of information. ggplot2 supplements boxplots with two geoms that show more information.

The first is geom_dotplot(). If you set the binaxis parameter of geom_dotplot() to "y", geom_dotplot() behaves like geom_boxplot(), display a separate distribution for each group of data.

Here each group functions like a vertical histogram. Add the parameter stackdir = "center" then re-run the code. Can you interpret the results?

```{r}
ggplot(data = mpg) +
  geom_dotplot(mapping = aes(x = class, y = hwy), binaxis = "y", 
               dotsize = 0.5, binwidth = 1)
```

```{r}
ggplot(data = mpg) +
  geom_dotplot(mapping = aes(x = class, y = hwy), binaxis = "y", 
               dotsize = 0.5, binwidth = 1, stackdir = "center")
```


geom_violin()

geom_violin() provides a second alternative to geom_boxplot(). A violin plot uses densities to draw a smoothed version of the centered dotplot you just made.

You can think of a violin plot as an outline drawn around the edges of a centered dotplot. Each “violin” spans the range of the data. The violin is thick where there are many values, and thin where there are few.

Convert the plot below from a boxplot to a violin plot. Note that violin plots do not use the parameters you saw for dotplots.

```{r}
ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy))

```


```{r}
ggplot(data = mpg) +
  geom_violin(mapping = aes(x = class, y = hwy))
```


```{r}
ggplot(data = mpg) +
  geom_violin(mapping = aes(x = class, y = hwy), draw_quantiles =  c(0.25, 0.5, 0.75))
```


geom_count()

Boxplots provide an efficient way to explore the interaction of a continuous variable and a categorical variable. But what if you have two categorical variables?

You can see how observations are distributed across two categorical variables with geom_count(). geom_count() draws a point at each combination of values from the two variables. The size of the point is mapped to the number of observations with this combination of values. Rare combinations will have small points, frequent combinations will have large points.


```{r}
ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = clarity))
```

Heat maps

Heat maps provide a second way to visualize the relationship between two categorical variables. They work like count plots, but use a fill color instead of a point size, to display the number of observations in each combination.


```{r}
diamonds %>% 
   count(cut, clarity) %>% 
   ggplot() +
     geom_tile(mapping = aes(x = cut, y = clarity, fill = n))
```

reproducir el siguiente heatmap
```{r}
diamonds %>% count(cut, color) %>% ggplot()+ geom_tile(aes(x=color, y=cut, fill=n))

```

# Boxplot Recap 

Recap

Boxplots, dotplots and violin plots provide an easy way to look for relationships between a continuous variable and a categorical variable. Violin plots convey a lot of information quickly, but boxplots have a head start in popularity — they were easy to use when statisticians had to draw graphs by hand.

In any of these graphs, look for distributions, ranges, medians, skewness or anything else that catches your eye to change in an unusual way from distribution to distribution. Often, you can make patterns even more revealing with the fct_reorder() function from the forcats package (we’ll wait to learn about forcats until after you study factors).
Count plots and heat maps help you see how observations are distributed across the interactions of two categorical variables. 


---
# Scatterplots

A scatterplot displays the relationship between two continuous variables. Scatterplots are one of the most common types of graphs—in fact, you’ve met scatterplots already in Visualization Basics.

In this tutorial, you’ll learn how to:

    Make new types of scatterplots with geom_text() and geom_jitter()
    Add multiple layers of geoms to a plot
    Enhance scatterplots with geom_smooth(), geom_rug(), and geom_repel()
    Change the coordinate system of a plot

The tutorial is adapted from R for Data Science by Hadley Wickham and Garrett Grolemund, published by O’Reilly Media, Inc., 2016, ISBN: 9781491910399. You can purchase the book at shop.oreilly.com.

The tutorial uses the ggplot2, ggrepel, and dplyr packages, which have been pre-loaded for your convenience.


## Scatterplots
Review 1 - geom_point()

In Visualization Basics, you learned how to make a scatterplot with geom_point().

The code below summarises the mpg data set and begins to plot the results. Finish the plot with geom_point(). Put mean_cty on the x
axis and mean_hwy on the y axis.


```{r}
mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot()+
  geom_point(aes(x=mean_cty, y=mean_hwy))
```

geom_text()

geom_text() and geom_label() create scatterplots that use words instead of points to display data. Each requires the extra aesthetic label, which you should map to a variable that contains text to display for each observation.

Convert the plot below from geom_point() to geom_text() and map the label aesthetic to the class variable. When you are finished convert the code to geom_label() and rerun the plot. Can you spot the difference?


```{r}
mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot() +
    geom_text(mapping = aes(x = mean_cty, y = mean_hwy, label = class))

mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot() +
    geom_label(mapping = aes(x = mean_cty, y = mean_hwy, label = class))
    
```

geom_smooth()

In Visualization Basics, you met geom_smooth(), which provides a summarised version of a scatterplot.

geom_smooth() uses a model to fit a smoothed line to the data and then visualizes the results. By default, geom_smooth() fits a loess smooth to data sets with less than 1,000 observations, and a generalized additive model to data sets with more than 1,000 observations.

method

You can use the method parameter of geom_smooth() to fit and display other types of model lines. To do this, pass method the name of an R modeling function for geom_smooth() to use, such as lm (for linear models) or glm (for generalized linear models).

In the code below, use geom_smooth() to draw the linear model line that fits the data.

```{r}
mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot() +
    geom_smooth(mapping = aes(x = mean_cty, y = mean_hwy), method = lm)
```

Layers
Add a layer

geom_smooth() becomes much more useful when you combine it with geom_point() to create a scatterplot that contains both:

    raw data
    a trend line

In ggplot2, you can add multiple geoms to a plot by adding multiple geom functions to the plot call. For example, the code below creates a plot that contains both points and a smooth line. Imagine what the results will look like in your head, and then run the code to see if you are right.


```{r}
mpg %>% 
  group_by(class) %>% 
  summarise(mean_cty = mean(cty), mean_hwy = mean(hwy)) %>% 
  ggplot() +
    geom_point(mapping = aes(x = mean_cty, y = mean_hwy)) +
    geom_smooth(mapping = aes(x = mean_cty, y = mean_hwy), method = lm) 
```

